---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# pkgcache

> Cache CRAN-like metadata and package files

[![lifecycle](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/pkgcache)](https://cran.r-project.org/package=pkgcache)
[![Travis build status](https://travis-ci.org/r-lib/pkgcache.svg?branch=master)](https://travis-ci.org/r-lib/pkgcache)
[![AppVeyor build status](https://ci.appveyor.com/api/projects/status/github/r-lib/pkgcache?branch=master&svg=true)](https://ci.appveyor.com/project/r-lib/pkgcache)
[![Coverage status](https://codecov.io/gh/r-lib/pkgcache/branch/master/graph/badge.svg)](https://codecov.io/github/r-lib/pkgcache?branch=master)

Metadata and package cache for CRAN-like repositories. This is a utilitiy
package to be used by package management tools that want to take advantage
of caching.

## Installation

You can install the released version of pkgcache from
[CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("pkgcache")
```

## Metadata cache

`cran_list()` lists all packages in the metadata cache. It includes
BioConductor package, and all versions (i.e. both binary and source)
of the packages for the current platform and R version.

```{r}
library(pkgcache)
cran_list()
```

`cran_deps()` and `cran_revdeps()` can be used to look up dependencies
and reverse dependencies:

```{r}
cran_deps("crayon")
```

```{r}
cran_revdeps("crayon")
```

The metadata is updated automatically if it is older than seven days,
and it can also be updated manually with `cran_update()`.

See the `cranlike_metadata_cache` R6 class for a lower level API, including
asynchronous methods, and more control.

## Package cache

Package management tools may use the `cache_*` functions and in particular
the `package_cache` class, to make use of local caching of package files.

The `cache_*` API is high level, and uses a user level cache:

```{r}
cache_summary()
```

```{r}
cache_list()
```

```{r}
cache_find(package = "dplyr")
```

`cache_add_file()` can be used to add a file, `cache_delete_files()` to
remove files, `cache_get_files()` to copy files out of the cache.

The `package_cache` class provides a finer API, including asynchronous
methods.

## Code of Conduct

Please note that the 'pkgcache' project is released with a
[Contributor Code of Conduct](.github/CODE_OF_CONDUCT.md). By contributing
to this project, you agree to abide by its terms.

## License

MIT © [RStudio Inc](https://rstudio.com)
