
```{r setup2, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  cache = TRUE
)
options(cli.unicode = FALSE)
```

Metadata and package cache for CRAN-like repositories. This is a utility
package to be used by package management tools that want to take advantage
of caching.

## Installation

You can install the released version of pkgcache from
[CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("pkgcache")
```

## Metadata cache

`meta_cache_list()` lists all packages in the metadata cache. It includes
Bioconductor package, and all versions (i.e. both binary and source)
of the packages for the current platform and R version.

```{r}
library(pkgcache)
meta_cache_list()
```

`meta_cache_deps()` and `meta_cache_revdeps()` can be used to look up
dependencies and reverse dependencies.

The metadata is updated automatically if it is older than seven days,
and it can also be updated manually with `meta_cache_update()`.

See the `cranlike_metadata_cache` R6 class for a lower level API,
and more control.

## Package cache

Package management tools may use the `pkg_cache_*` functions and in
particular the `package_cache` class, to make use of local caching of
package files.

The `pkg_cache_*` API is high level, and uses a user level cache:

```{r}
pkg_cache_summary()
```

```{r}
pkg_cache_list()
```

```{r}
pkg_cache_find(package = "dplyr")
```

`pkg_cache_add_file()` can be used to add a file,
`pkg_cache_delete_files()` to remove files, `pkg_cache_get_files()` to
copy files out of the cache.

The `package_cache` class provides a finer API.

## Bioconductor support

Both the metadata cache and the package cache support Bioconductor by
default, automatically. See the `BioC_mirror` option and the `R_BIOC_MIRROR`
and `R_BIOC_VERSION` environment variables below to configure pkgcache's
Bioconductor support.

## Package Options

- The `BioC_mirror` option can be used to select a Bioconductor mirror.
  This takes priority over the `R_BIOC_MIRROR` environment variable.
- `pkgcache_timeout` is the HTTP timeout for all downloads. It is in
  seconds, and the limit for downloading the whole file. Defaults to 3600,
  one hour. It corresponds to the [`TIMEOUT` libcurl option](https://curl.haxx.se/libcurl/c/CURLOPT_TIMEOUT.html).
- `pkgcache_connecttimeout` is the HTTP timeout for the connection phase.
  It is in seconds and defaults to 30 seconds. It corresponds to the
  [`CONNECTTIMEOUT` libcurl option](https://curl.haxx.se/libcurl/c/CURLOPT_CONNECTTIMEOUT.html).
- `pkgcache_low_speed_limit` and `pkgcache_low_speed_time` are used for a
  more sensible HTTP timeout. If the download speed is less than
  `pkgcache_low_speed_limit` bytes per second for at least
  `pkgcache_low_speed_time` seconds, the download errors. They correspond
  to the [`LOW_SPEED_LIMIT`](https://curl.haxx.se/libcurl/c/CURLOPT_LOW_SPEED_LIMIT.html) and
  [`LOW_SPEED_TIME`](https://curl.haxx.se/libcurl/c/CURLOPT_LOW_SPEED_TIME.html)
  curl options.

## Package environment variables

- The `R_BIOC_VERSION` environment variable can be used to override the
  default Bioconductor version detection and force a given version. E.g.
  this can be used to force the development version of Bioconductor.
- The `R_BIOC_MIRROR` environment variable can be used to select a
  Bioconductor mirror. The `BioC_mirror` option takes priority over this,
  if set.
- `PKGCACHE_TIMEOUT` is the HTTP timeout for all downloads. It is in
  seconds, and the limit for downloading the whole file. Defaults to 3600,
  one hour. It corresponds to the [`TIMEOUT` libcurl option](https://curl.haxx.se/libcurl/c/CURLOPT_TIMEOUT.html).
  The `pkgcache_timeout` option has priority over this, if set.
- `PKGCACHE_CONNECTTIMEOUT` is the HTTP timeout for the connection phase.
  It is in seconds and defaults to 30 seconds. It corresponds to the
  [`CONNECTTIMEOUT` libcurl option](https://curl.haxx.se/libcurl/c/CURLOPT_CONNECTTIMEOUT.html). The
  `pkgcache_connecttimeout` option takes precedence over this, if set.
- `PKGCACHE_LOW_SPEED_LIMIT` and `PKGCACHE_LOW_SPEED_TIME` are used for a
  more sensible HTTP timeout. If the download speed is less than
  `PKGCACHE_LOW_SPEED_LIMIT` bytes per second for at least
  `PKGCACHE_LOW_SPEED_TIME` seconds, the download errors. They correspond
  to the [`LOW_SPEED_LIMIT`](https://curl.haxx.se/libcurl/c/CURLOPT_LOW_SPEED_LIMIT.html) and
  [`LOW_SPEED_TIME`](https://curl.haxx.se/libcurl/c/CURLOPT_LOW_SPEED_TIME.html)
  curl options. The `pkgcache_low_speed_time` and `pkgcache_low_speed_limit`
  options have priority over these environment variables, if they are set.

## Code of Conduct

Please note that the 'pkgcache' project is released with a
[Contributor Code of Conduct](https://github.com/r-lib/pkgcache/blob/master/.github/CODE_OF_CONDUCT.md).
By contributing to this project, you agree to abide by its terms.

## License

MIT © [RStudio Inc](https://rstudio.com)
