
```{r setup2, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  cache = TRUE
)
options(cli.unicode = FALSE)
options(cli.dynamic = FALSE)
options(crayon.enabled = FALSE)
crayon::num_colors(forget=TRUE)
```

Metadata and package cache for CRAN-like repositories. This is a utility
package to be used by package management tools that want to take advantage
of caching.

## Installation

You can install the released version of pkgcache from
[CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("pkgcache")
```

If you need the development version, you can install it from
[GitHub](https://github.com) with:

``` r
pak::pak("r-lib/pkgcache")
```

## Metadata cache

`meta_cache_list()` lists all packages in the metadata cache. It includes
Bioconductor package, and all versions (i.e. both binary and source)
of the packages for the current platform and R version.

```{r include = FALSE}
# If we don't do this, then Unicode characters end up in documentation
# files.
library(pkgcache)
meta_cache_list()
```

(We load the pillar package, because it makes the pkgcache data frames
print nicer, similarly to tibbles.)

```{r}
library(pkgcache)
library(pillar)
meta_cache_list()
```

`meta_cache_deps()` and `meta_cache_revdeps()` can be used to look up
dependencies and reverse dependencies.

The metadata is updated automatically if it is older than seven days,
and it can also be updated manually with `meta_cache_update()`.

See the `cranlike_metadata_cache` R6 class for a lower level API,
and more control.

## Package cache

Package management tools may use the `pkg_cache_*` functions and in
particular the `package_cache` class, to make use of local caching of
package files.

The `pkg_cache_*` API is high level, and uses a user level cache:

```{r}
pkg_cache_summary()
```

```{r}
pkg_cache_list()
```

```{r}
pkg_cache_find(package = "dplyr")
```

`pkg_cache_add_file()` can be used to add a file,
`pkg_cache_delete_files()` to remove files, `pkg_cache_get_files()` to
copy files out of the cache.

The `package_cache` class provides a finer API.

## Installed packages

pkgcache contains a very fast DCF parser to parse `PACKAGES*` files,
or the `DESCRIPTION` files in installed packages.
`parse_packages()` parses all fields from `PACKAGES`, `PACKAGES.gz` or
`PACKAGES.rds` files.
`parse_installed()` reads
_all_ metadata from packages installed into a library:

```{r}
parse_installed()
```

## Bioconductor support

Both the metadata cache and the package cache support Bioconductor by
default, automatically. See the `BioC_mirror` option and the `R_BIOC_MIRROR`
and `R_BIOC_VERSION` environment variables below to configure
Bioconductor support.

## Package Options

- The `BioC_mirror` option can be used to select a Bioconductor mirror.
  This takes priority over the `R_BIOC_MIRROR` environment variable.
- You can use the `pkg.current_platform` option to set the platform string
  for the current platform for the `current_r_platform()` function. This is
  useful if pkgcache didn't detect the platform correctly. Alternatively,
  you can use the `PKG_CURRENT_PLATFORM` environment variable. The option
  takes priority.
- `pkgcache_timeout` is the HTTP timeout for all downloads. It is in
  seconds, and the limit for downloading the whole file. Defaults to 3600,
  one hour. It corresponds to the [`TIMEOUT` libcurl option](https://curl.se/libcurl/c/CURLOPT_TIMEOUT.html).
- `pkgcache_connecttimeout` is the HTTP timeout for the connection phase.
  It is in seconds and defaults to 30 seconds. It corresponds to the
  [`CONNECTTIMEOUT` libcurl option](https://curl.se/libcurl/c/CURLOPT_CONNECTTIMEOUT.html).
- `pkgcache_low_speed_limit` and `pkgcache_low_speed_time` are used for a
  more sensible HTTP timeout. If the download speed is less than
  `pkgcache_low_speed_limit` bytes per second for at least
  `pkgcache_low_speed_time` seconds, the download errors. They correspond
  to the [`LOW_SPEED_LIMIT`](https://curl.se/libcurl/c/CURLOPT_LOW_SPEED_LIMIT.html) and
  [`LOW_SPEED_TIME`](https://curl.se/libcurl/c/CURLOPT_LOW_SPEED_TIME.html)
  curl options.

## Package environment variables

- The `R_BIOC_VERSION` environment variable can be used to override the
  default Bioconductor version detection and force a given version. E.g.
  this can be used to force the development version of Bioconductor.
- The `R_BIOC_MIRROR` environment variable can be used to select a
  Bioconductor mirror. The `BioC_mirror` option takes priority over this,
  if set.
- You can use the `PKG_CURRENT_PLATFORM` environment variable to set the
  platform string for the current platform for the `current_r_platform()`
  function. This is useful if pkgcache didn't detect the platform correctly.
  Alternatively, you can use the `pkg.current_platofrm` option, which takes.
  priority over the environment variable.
- `PKGCACHE_PPM_REPO` is the name of the Posit Package Manager repository
  to use. Defaults to `"cran"`.
- `PKGCACHE_PPM_URL` is the base URL of the Posit Package Manager instance
  to use. It defaults to the URL of the Posit Public Package Manager
  instance at <https://packagemanager.posit.co/client/#/>.
- `PKGCACHE_TIMEOUT` is the HTTP timeout for all downloads. It is in
  seconds, and the limit for downloading the whole file. Defaults to 3600,
  one hour. It corresponds to the [`TIMEOUT` libcurl option](https://curl.se/libcurl/c/CURLOPT_TIMEOUT.html).
  The `pkgcache_timeout` option has priority over this, if set.
- `PKGCACHE_CONNECTTIMEOUT` is the HTTP timeout for the connection phase.
  It is in seconds and defaults to 30 seconds. It corresponds to the
  [`CONNECTTIMEOUT` libcurl option](https://curl.se/libcurl/c/CURLOPT_CONNECTTIMEOUT.html). The
  `pkgcache_connecttimeout` option takes precedence over this, if set.
- `PKGCACHE_LOW_SPEED_LIMIT` and `PKGCACHE_LOW_SPEED_TIME` are used for a
  more sensible HTTP timeout. If the download speed is less than
  `PKGCACHE_LOW_SPEED_LIMIT` bytes per second for at least
  `PKGCACHE_LOW_SPEED_TIME` seconds, the download errors. They correspond
  to the [`LOW_SPEED_LIMIT`](https://curl.se/libcurl/c/CURLOPT_LOW_SPEED_LIMIT.html) and
  [`LOW_SPEED_TIME`](https://curl.se/libcurl/c/CURLOPT_LOW_SPEED_TIME.html)
  curl options. The `pkgcache_low_speed_time` and `pkgcache_low_speed_limit`
  options have priority over these environment variables, if they are set.
- `R_PKG_CACHE_DIR` is used for the cache directory, if set. (Otherwise
  `rappdirs::user_cache_dir()` is used, see also `meta_cache_summary()` and
  `pkg_cache_summary()`).

## Using pkgcache in CRAN packages

If you use pkgcache in your CRAN package, please make sure that

- you don't use pkgcache in your examples, and
- you set the `R_USER_CACHE_DIR` environment variable to a temporary
  directory (e.g. via `tempfile()`) during test cases. See the
  `tests/testthat/setup.R` file in pkgcache for an example.

This is to make sure that pkgcache does not modify the user's files while
running `R CMD check`.

## Code of Conduct

Please note that the pkgcache project is released with a
[Contributor Code of Conduct](https://r-lib.github.io/pkgcache/CODE_OF_CONDUCT.html).
By contributing to this project, you agree to abide by its terms.

## License

MIT (c) [Posit Software, PBC](https://posit.co)
